---
title: "phyloseq"
output: pdf_document
date: "2024-04-18"
---

#Install Phyloseq
```{r}
if (!require("BiocManager", quietly = TRUE))
  install.packages("BiocManager")
```
```{r}
if (!require("BiocManager", quietly = TRUE))
  install.packages("BiocManager")

BiocManager::install("phyloseq")
```


BiocManager::install("phyloseq")

# Install packages
```{r}
source('http://bioconductor.org/biocLite.R')
biocLite('phyloseq')
```
```{r}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("phyloseq")
```
```{r}
library(phyloseq)
```

```{r}
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("phyloseq")
```
# Load required packages
```{r}
library(phyloseq)
```

```{r}
library(dplyr)
```

```{r}
library(BiMiCo)
```

# Load taxa and seqtab.nochim

```{r}
load("RData/taxa.RData")
load("RData/seqtab.nochim.RData")
```

# import metadata
```{r}
meta<- read.csv("metadata.csv" , header=TRUE, row.names = 1) 
```

# create phyloseq object
```{r}
#make sure the seqtab.nochim and taxa objects are loaded
physeq <- phyloseq(otu_table(seqtab.nochim, taxa_are_rows=FALSE), sample_data(metadata), tax_table(taxa))
physeq
```
                     
#Transform sample counts
```{r}
# convert from raw to relative abundance
physeq <- transform_sample_counts(physeq, function(abund) 1*(abund>0))
physeq
```

# Remove the sequence itself and replace with ASV
```{r}
dna <- Biostrings::DNAStringSet(taxa_names(phyloseq))
names(dna) <- taxa_names(physeq)
physeq <- merge_phyloseq(physeq, dna)
taxa_names(physeq) <- paste0("ASV", seq(ntaxa(physeq)))
physeq
```
# Remove mitochondria and chloroplast matches. Remove all non bacterial sequences
```{r}
physeq <- physeq %>% subset_taxa( Family!= "Mitochondria" | is.na(Family) & Order!="Chloroplast" | is.na(Order) )
physeq
```

# Install BiMiCo
```{r}
install.packages("remotes")
remotes::install_github("peterolah001/BiMiCo")
```

# Remove all non bacterial sequences
```{r}
physeq<-rm_nonbac(physeq)
physeq
```

# Save physeq objects to load later
```{r}
save(physeq, file="RData/physeq.RData")
```

# Pushing to GitHub
# Open Ubuntu
# Navigate to bioinformatics folder (insert your own user name after /Users/)
cd /mnt/c/Users/eh040/Desktop/Bioinformatics
git status
# To upload all of the MicrobiomeDADA2 files, type in "git add filename". You'll have to do each one individually
# Once previous step is completed:
git status
# All the files you want to upload should be listed in green
git commit -m "phyloseq"
git push




